Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{DISTRIBUTION STATEMENT A. Approved for public release.}

### 2022-02-24 paper/agree.tex
Full pass done. Spell check. Minor edits. Paper fits.
snapshot

The goal of this section is to illustrate in more detail the process a
system designer follows to add cyber requirements to the AGREE
specifications and then automatically transform the system to insert
the high-assurance components. The AGREE specifications generated by
the transforms for the high-assurance components are
explained. The section ends with a concise formal statement of the
meaning of an AGREE specification for a high-assurance component. This
meaning is what must be preserved by the synthesis.

@@

The AGREE specification for the SW component in the example of Section~\ref{sec:example} with the added cyber requirements is given in \figref{fig:sw}. The AGREE specification language is a first-order predicate calculus that uses stream concepts, and operators, from the Lustre language \cite{10.1145/41625.41641}. As with Lustre, the semantics are synchronous data-flow where the inputs, outputs, and expressions are characterized by data streams that comply with the input assumptions. The semantics are such that the contracts are evaluated in dependency order with inputs being propagated to outputs through all the contracts until they stabilize; as such, the contracts, and thereby the top-level model, must be acyclic. Once the contracts have stabilized, the model takes a synchronous step to the next input data in the stream. The semantics do not model computation or communication delay. The output of one contract is seen at the input of any downstream contract in the same step of the input data stream.

The AGREE model checker attempts to prove several properties of the top-level model being verified. The first is that the output guarantees of each component implementing the system are strong enough to validate the input assumptions of any downstream component as well as to satisfy the guarantees of the output of the top-level component being verified (i.e., the system composition meets input assumptions at each input as well as the guarantees on the system output). These properties are reported in the expanded lists in \figref{fig:example}(b) and \figref{fig:hardened}(b).  The next set of properties prove that the contract specifications for each component are self-consistent (i.e, a contract does not contradict itself). These are the unexpanded results at the bottom of the figures.

Returning back to the contract in \figref{fig:sw}, it uses \texttt{eq} statements to define variables local to the contract specification. For example, the \texttt{req} variable is equivalent to the \texttt{event(AutomationRequest)} expression. In the AGREE semantics, there is an implicit \emph{event} input (or output) associated with every named event port in a component. The semantics used here do not buffer these events so the implicit input (or output) is only a boolean value. An \texttt{event} expression refers to that implicit input (or output) and is true when data is placed on the named port. The system contract here states assumptions on well-formed input, followed by guarantees on properties about the output.

@@

\subsection{Brief Semantics Definition}
\label{agree-semantics}
Here the formal semantics of the AGREE contract specification are briefly
presented to make clear the meaning of a high-assurance component.
These semantics are used in the next section to argue that the synthesis
preserves the same input to output behavior as the AGREE specification.

Assume that all data is in its raw form which is a contiguous sequence of bytes
representing exactly what is sent over a wire by the communication fabric, so a
datum is given by a \konst{string}. This assumption is important to the
correctness argument in the next section. An \emph{environment},
$\theta: \lval \mapsto \konst{string}$, binds \emph{L-values} to strings where
an L-value is anything that can appear on the left-hand side of an assignment
such as a named port, a field in a record, a entry in an array, a local
value defined by an eq-statement etc.

>>>>

The goal of this section is to illustrate in more detail the process a
system designer follows to add cyber requirements to the AGREE
specifications and then automatically transform the system to insert
the high-assurance components. The AGREE specifications generated by
the transforms for the high-assurance components are also
explained. The section ends with a concise formal statement of the
meaning of an AGREE specification for a high-assurance component. This
meaning is what must be preserved by the synthesis.

@@

The AGREE specification for the SW component in the example of
Section~\ref{sec:example} with the added cyber requirements is given
in \figref{fig:sw}. The AGREE specification language is a first-order
predicate calculus that uses stream concepts, and operators, from the
Lustre language \cite{10.1145/41625.41641}. As with Lustre, the
semantics are synchronous dataflow where the inputs, outputs, and
expressions are data streams that comply with the input
assumptions. Contracts are evaluated in dependency order with inputs
being propagated to outputs through all contracts until they
stabilize; as such, the contracts, and thereby the top-level model,
must be acyclic.\footnote{An apparent syntactic cycle, where a
component is linked back to itself, may be broken temporally by
inserting delay elements.} Once the contracts have stabilized, the
model takes a synchronous step to the next input data in the
stream. The semantics do not model computation or communication
delay. The output of one contract is seen at the input of any
downstream contract in the same step of the input data stream.

The AGREE model checker attempts to prove several properties of the
top-level model being verified. The first is that the output
guarantees of each component implementing the system are strong enough
to validate the input assumptions of any downstream component as well
as to satisfy the guarantees of the output of the top-level component
being verified (\ie, the system composition meets input assumptions
at each input as well as the guarantees on the system output). These
properties are reported in the expanded lists
in \figref{fig:example}(b) and \figref{fig:hardened}(b).  The next set
of properties prove that the contract specifications for each
component are \emph{self-consistent} \ie, a contract does not contradict
itself. These are the unexpanded results at the bottom of the
figures.

Returning back to the contract of \figref{fig:sw}, it uses \texttt{eq}
statements to define variables local to the contract
specification. For example, the \texttt{req} variable is equivalent to
the \texttt{event(AutomationRequest)} expression. In the AGREE
semantics, there is an implicit \emph{event} input (or output)
associated with every named event port in a component. The semantics
used here do not buffer these events so the implicit input (or output)
is only a boolean value. An \texttt{event} expression refers to that
implicit input (or output) and is true when data is placed on the
named port. The system contract here states assumptions on well-formed
input, followed by guarantees on properties about the output.

@@

\subsection{AGREE Semantics}
\label{agree-semantics}

Here the formal semantics of the AGREE contract specification are
briefly presented to make clear the meaning of a high-assurance
component.  These semantics are the basis of proofs showing that a
synthesized component has input to output behavior specified by its
contract.

Assume that all data flowing between components is in the form of byte
arrays, thus input and output is modelled at the level of what is
passed ``over the wire'' by the communication fabric. An \emph{environment},
$\theta: \lval \mapsto \konst{string}$, binds \emph{L-values} to
strings where an L-value is anything that can appear on the left-hand
side of an assignment such as a named port, a field in a record, a
entry in an array, a local value defined by an eq-statement etc.
